---
title: "A1_Individual Project"
author: "Justine Franco"
date: "2023-07-09"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Weekly sales forecasting starter:
# Install forecast-package first.
library(tidyverse)
library(forecast)
library(shiny)

# historical records from multiple stores
fsales <- "https://raw.githubusercontent.com/multidis/hult-inter-bus-reports-r/main/forecasting/sales_weekly.csv"
sales <- read_csv(fsales)
sales

# latest (current) week
nweek_now <- max(sales$Week)

# calculate the range of stores
store_range <- range(sales$Store)
store_min <- store_range[1]
store_max <- store_range[2]

# calculate the range of weeks
week_range <- range(sales$Week)
week_min <- week_range[1]
week_max <- week_range[2]

# add year, month, and quarter to the sales table
sales$Year = floor((sales$Week - 1) / 52) + 1

# Calculate Month
sales$Month = ceiling(((sales$Week - 1) %% 52) / 4.33) + 1

# Calculate Quarter
sales$Quarter = ceiling(sales$Month / 3)

# find max year
year_max <- max(sales$Year)

# find the max month of the current year
max_month_currYear <- sales %>%
  filter(Year == year_max) %>%
  pull(Month) %>%
  max()

```

I am an analyst of a famous retail brand. I performed sales forecasting for upcoming time periods to help my colleagues to explore the forecasts. Here is my web application:

## Analyse 1: Provide a current week performance indicator in the UI showing current week's actual sales vs. forecast for the current week computed at the end of the last week.

In this first analysis, I provide a performance indicator in the UI that compares the actual sales for the current week with the forecast computed at the end of the last week. The actual sales and forecast data are filtered based on the current week. 

The performance indicator is calculated by comparing the actual sales value with the forecast value for the current week. This provides insights into how well the sales forecast performed for the current week. The performance indicator is then displayed in the UI to provide a snapshot of the current week's performance relative to the forecast.

```{r, echo=FALSE}
library(shiny)
library(forecast)

inputPanel(selectInput(
  "storenum",
  label = "Select store: ",
  choices = 1:max(sales$Store),
  selected = 1
),)

# latest (current) week
nweek_now <- max(sales$Week)

renderDataTable({
  # check forecast accuracy for the most recent quarter
  sales_hist <- sales %>%
    filter(Store == input$storenum) %>%
    subset(Week < nweek_now)
  sales_last <- sales %>%
    filter(Store == input$storenum) %>%
    subset(Week >= nweek_now)
  
  # time series with annual periodicity to account seasonality
  sales_hist_ts <- ts(sales_hist$Weekly_Sales, frequency = 52)
  # autoplot(sales_hist_ts)
  
  # ARIMA: Auto-Regressive Integrated Moving Average
  # methodological details:
  #  https://otexts.com/fpp3/arima.html
  arima_model <- auto.arima(sales_hist_ts, seasonal.test = "seas")
  
  # forecast horizon to match most recent quarter
  arima_pred <- forecast(arima_model, h = 1)
  
  # note: confidence intervals (lower, upper) are available as well
  sales_pred_eval <-
    data.frame(
      predicted = as.numeric(arima_pred$mean),
      actual = sales_last$Weekly_Sales,
      Week = sales_last$Week
    )
  sales_pred_eval
  
})

```

## Analyse 2: Generate sales forecast for a specific future quarter (during the upcoming year) selected in the UI.

In this analysis, the goal is to generate a sales forecast for a specific future quarter, which is selected in the user interface (UI). The UI provides two input panels: one to select the store and another to select the quarter.

The code generates a graph using ggplot2 to visualize the forecasted sales for the selected quarter. The x-axis represents the months, and the y-axis represents the "Sales Forecast".

This analysis allows users to select a specific store and quarter to generate sales forecasts, providing valuable insights for planning and decision-making.

```{r, echo=FALSE}
library(shiny)
library(forecast)
library(tidyverse)
library(plotly)

# Define store_min and store_max
store_min <- 1
store_max <- 45

# Input Panels
inputPanel(
  selectInput(
    "store_1",
    label = "Select Store: ",
    choices = 1:max(sales$Store),
    selected = 1
  ),

  selectInput(
    "quarter_1",
    label = "Select Quarter: ",
    choices = c(
      "Q1: Jan-Mar" = 1, 
      "Q2: Apr-Jun" = 2, 
      "Q3: Jul-Sep" = 3, 
      "Q4: Oct-Dec" = 4
    ),
    selected = 1
  )
)

renderPlot({
  # Calculate the monthly sums and filter the store to analyze the quarters
  sales_1 <- sales %>%
    filter(Store == input$store_1) %>%
    group_by(Year, Quarter, Month) %>%
    summarize(Monthly_Sales = sum(Weekly_Sales))
  
  sales_hist_ts <- ts(sales_1$Monthly_Sales, frequency = 12)
  
  # methodological details:
  #  https://otexts.com/fpp3/arima.html
  # ARIMA: Auto-Regressive Integrated Moving Average
  arima_model <- auto.arima(sales_hist_ts, seasonal.test = "seas")
  
  # forecast horizon to match most recent quarter
  arima_pred <- forecast(arima_model, h = 12)
  
  # format the dataframe
  sales_pred_eval <-
    data.frame(Prediction = as.numeric(arima_pred$mean),
               Month = seq.int(1, 12))
  
  # add quarters to the months
  sales_pred_eval$Quarter <- ceiling(sales_pred_eval$Month / 3)
  
  # filter, to show only the selected quarter
  sales_pred_eval <- sales_pred_eval %>%
    filter(Quarter == input$quarter_1)
  
  # add adaptive labels
  quarter_text <- paste("Quarter", input$quarter_1)
  
  # define month names
  month_names <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                   "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  
  # plot chart
  ggplot() +
    geom_line(data = sales_pred_eval,
              aes(x = Month, y = Prediction, color = quarter_text)) +
    xlab("Months") +
    ylab("Sales Forecast") +
    scale_x_continuous(breaks = seq(1, 12), labels = month_names) +
    theme_bw()
})

```


## Analysis 3: Generate sales forecast for a given week (within the next three months) selected in the UI 

In this analysis, a Shiny application is created to generate a sales forecast for a given week over the next three months, based on the selected store.

The user interface consists of a slider for selecting the forecast horizon in weeks and an input for selecting the store. 

The sales forecast graph shows actual sales data for the selected store as a green line, and forecast sales for the coming weeks as a red line. The x-axis represents weeks and the y-axis represents sales. 

The table on the main panel displays forecast sales for each week of the forecast horizon. It includes the week number and the corresponding forecast sales value.

Overall, this analysis enables users to explore and visualize sales forecasts for a specific week into the future, providing valuable information for planning and decision-making.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(shiny)
library(forecast)

# Create the Shiny app
ui <- fluidPage(
  titlePanel("Forecast"),
    sidebarPanel(
      sliderInput("horizon", "Forecast Horizon (weeks)", min = 1, max = 12, value = 12),
      selectInput("store_3", "Select Store", choices = unique(sales$Store), selected = 1)
    ),
    mainPanel(
      plotlyOutput("salesChart"),
      tableOutput("salesTable")
    )
  )

server <- function(input, output) {
  # Generate the sales forecast chart
  output$salesChart <- renderPlotly({
    # Filter the sales data for the selected store
    sales_store <- sales %>% 
      filter(Store == input$store_3)
    
    # Perform forecasting for the selected store
    sales_ts <- ts(sales_store$Weekly_Sales, frequency = 52)
    
    # ARIMA: Auto-Regressive Integrated Moving Average
    # methodological details:
    #  https://otexts.com/fpp3/arima.html
    arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
    forecasted_sales <- forecast(arima_model, h = input$horizon)
    
    # Create the plotly object
    plot_ly() %>%
      add_lines(x = sales_store$Week, y = sales_store$Weekly_Sales, name = "Actual Sales", type = "scatter", mode = "lines", line = list(color = "green")) %>%
      add_lines(x = (max(sales_store$Week) + 1):(max(sales_store$Week) + input$horizon), y = forecasted_sales$mean, name = "Forecast", type = "scatter", mode = "lines", line = list(color = "red")) %>%
      layout(
        title = "Sales Forecast",
        xaxis = list(title = "Weeks"),
        yaxis = list(title = "Sales")
      )
  })
  
  # Generate the forecasted sales table
  output$salesTable <- renderTable({
    # Filter the sales data for the selected store
    sales_store <- sales %>% filter(Store == input$store_3)
    
    # Perform forecasting for the selected store
    sales_ts <- ts(sales_store$Weekly_Sales, frequency = 52)
    
    # ARIMA: Auto-Regressive Integrated Moving Average
    # methodological details:
    #  https://otexts.com/fpp3/arima.html
    arima_model <- auto.arima(sales_ts, seasonal.test = "seas")
    forecasted_sales <- forecast(arima_model, h = input$horizon)
    
    # Create a data frame with the forecasted sales
    forecast_table <- data.frame(Week = (max(sales_store$Week) + 1):(max(sales_store$Week) + input$horizon),
                                 Forecast = forecasted_sales$mean)
    
    forecast_table
  })
}

# Run the app
shinyApp(ui = ui, server = server)
```


## Analysis 4: Incorporate a UI functionality to compare monthly sales between the past year (actual) and the upcoming year (forecast).

The analysis focuses on incorporating a user interface (UI) functionality to compare monthly sales between the past year (actual) and the upcoming year (forecast). The Shiny app allows the user to select a specific store, start month, and end month through input elements in the sidebar.

Upon selecting the desired options, the server generates a sales forecast comparison plot. The plot compares the historical sales data from the current year (referred to as "Current Year" in green) with the forecasted sales data for the upcoming year (referred to as "Forecast Next Year" in red). The plot displays the sales data over weeks on the x-axis and the corresponding sales values on the y-axis.

Overall, this analysis enhances the forecasting capabilities by enabling users to compare monthly sales between the past and future periods in a dynamic and intuitive manner.

```{r, echo=FALSE}
library(shiny)
library(forecast)

# Define UI
ui <- fluidPage(
  titlePanel("Sales Forecast Comparison"),
  sidebarLayout(
    sidebarPanel(
      selectInput(
        "store_4",
        label = "Select Store:",
        choices = 1:max(sales$Store),
        selected = 1
      ),
      sliderInput(
        inputId = "slider_start",
        label = "Start Month:",
        min = 1,
        max = max_month_currYear,
        value = 1
      ),
      sliderInput(
        inputId = "slider_end",
        label = "End Month:",
        min = 1,
        max = max_month_currYear,
        value = max_month_currYear
      )
    ),
    mainPanel(
      plotOutput("salesPlot")
    )
  )
)

# Define server
server <- function(input, output) {
  # Generate the sales forecast comparison plot
  output$salesPlot <- renderPlot({
    # Filter sales data for the selected store
    sales_4 <- sales %>% filter(Store == input$store_4)
    
    # Calculate time series
    sales_hist_ts <- ts(sales_4$Weekly_Sales, frequency = 52)
    
    # ARIMA: Auto-Regressive Integrated Moving Average
    # methodological details:
    #  https://otexts.com/fpp3/arima.html
    arima_model <- auto.arima(sales_hist_ts, seasonal.test = "seas")
    
    # forecast horizon
    arima_pred <- forecast(arima_model, h = 52)
    
    # Format data frame with forecasted sales
    sales_pred_eval <- data.frame(
      Prediction = as.numeric(arima_pred$mean),
      Week = seq.int(max(sales_4$Week) + 1, max(sales_4$Week) + 52)
    )
    
    # Add Year and Month to data frame
    sales_pred_eval$Year <- floor((sales_pred_eval$Week) / 52) + 1
    sales_pred_eval$Month <- floor((sales_pred_eval$Week - (sales_pred_eval$Year - 1) * 52) / (52 / 12)) + 1
    
    # Filter months selected by sliders
    sales_pred_eval <- sales_pred_eval %>%
      filter(Month >= input$slider_start) %>%
      filter(Month <= input$slider_end)
    
    # Add week since the start of the year to compare with forecast
    sales_pred_eval$year_week <- sales_pred_eval$Week - (sales_pred_eval$Year - 1) * 52
    
    # Filter months selected by sliders for historical sales
    hist_sales <- sales_4 %>%
      filter(Year == year_max) %>%
      filter(Month >= input$slider_start) %>%
      filter(Month <= input$slider_end)
    
    # Add week since the start of the year to compare with forecast
    hist_sales$year_week <- hist_sales$Week - (hist_sales$Year - 1) * 52
    
    # Plot forecast and the current year based on filters
    ggplot() +
      geom_line(
        data = sales_pred_eval,
        aes(x = year_week, y = Prediction, color = "Forecast Next Year")
      ) +
      geom_line(
        data = hist_sales,
        aes(x = year_week, y = Weekly_Sales, color = "Current Year")
      ) +
      xlab("Weeks") +
      ylab("Sales") +
      scale_color_manual(
        values = c("Current Year" = "green", "Forecast Next Year" = "red")
      ) +
      labs(color = '') +
      theme_bw()
  })
}

# Run the app
shinyApp(ui = ui, server = server)
```

## Conclusion

In conclusion, I developed a customized web application for forecasting product demand in the retail sector. The web application allows users to explore historical sales variations, forecast sales for specific future quarters or selected weeks, and compare monthly sales between last year and the coming year. By integrating these features and providing interactive graphs and results, the web application enables users to make data-driven decisions and gain valuable insights into product demand trends.